---
title: "Data Hub Authorization"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: true
---
```{r include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent()

library(DiagrammeR)
knitr::knit_engines$set(incgraph = function(options) {
  assign(options$name, function(i) {
  graph <- paste(options$code, collapse = "\n")
  x <- 0
  while (TRUE) {
    x <- x + 1
    placeholder <- paste0("#", x)
    if (grepl(placeholder, graph))
    graph <- gsub(placeholder, if (x > i) "[style=invis]" else "", graph)
    else
    break
  }
  grViz(graph, width=750, height=options$height)
  }, .GlobalEnv)
  ""
})

xaringanExtra::use_tile_view()
```

# Integration

```{r echo=FALSE}
grViz("
      digraph {
        compound=true
        User -> Superset [lhead=cluster_platform]
        subgraph cluster_platform {
          {
            rank=same
            ClickHouse
            Superset
            '...' [color=white]
          }
          ClickHouse -> Superset -> '...' [style=invis]
        }
      }
", width=750, height=300)
```
???
- platform integrates different existing components
--

- Authentication
- Authorization
???
- creates uniform authentication and authorization layer across everything

---

# Authorization

```{r echo=FALSE}
grViz("
      graph {
        'Project A'
        'Project B'
        'Project C'
      }
", width=750, height=300)
```
???
- components have different ideas about how their data can be segmented
- maybe
  - component A has its own policy language for fine-grained access
  - component B has "projects"
  - component C is a database with RLS
--
name: segmentation
layout: true

Organization / Realm / Zone / Space / ...

---
???
- common language: projects, a flat segmentation of the application's data
- uniform principle throughout the platform

---
layout: true

# Keycloak Authorization Services

```{incgraph name='kcas', height=350}
digraph {
  p[label="in group X (Policy)"] #3
  p -> Tom [style=invis]
  {rank=same;nodesep=1;Tom;Apple}
  Apple[label="Apple (Resource)"] #1
  Tom -> Apple [label="apple:eat (Scope)"] #2
}
```

---
`r kcas(0)`
???
- basic building blocks of Keycloak Authorization Services
- Tom = user that wants to do things
---
`r kcas(1)`
???
- Resources = things that can be accessed / used
---
`r kcas(2)`
???
- Scopes = what you can do with a resource
---
`r kcas(3)`
???
- why Tom? because he's in group X

--
= Permission
???
- Permission = If **Policy** applies then allow **Scope** on **Resource**
---
layout: true

# Data Hub Resources

```{incgraph name='dhres', height=450}
graph {
  r[label=""] #3
  t1[label="Tenant 1"] #1
  t2[label="Tenant 2"] #1
  p1[label='Project A']
  p2[label='Project B'] #1
  p3[label='Project C'] #1
  c[label='Webhook Credential']
  p2 -- c #2
  g[label='Group Foobar']
  t2 -- g #2

  r -- t1 #3
  r -- t2 #3
  t1 -- p1 #2
  t1 -- p2 #2
  t2 -- p3 #2
}
```

---
`r dhres(0)`
???
- resources in Data Hub
- creating permissions for every single thing unfeasible
---
`r dhres(1)`
???
- of course more
- tenants are also resources
---
`r dhres(2)`
???
- resources can be organized in a tree
---
`r dhres(3)`
???
- at least when you turn the forest into a tree
- this simplifies things
  - can give a single permission for a whole subtree
  - recursive deletion instead of forgetting some credential when deleting a project etc.
- tree concept **on top** of KC AS because it's helpful
- same principles on all levels, allows expression of desired access in both directions
  - give admin permission, but for only one project
  - second tenant: admins only have access inside their tenant
  - contrast with simple admin true/false (either you're admin or regular user), does not solve either of those
---
layout: false
template: segmentation

???
- when to create S3 bucket / DB tables / dashboard space?
- manual = same chores every time
- same as arranging book pages in a book instead of loose in a bag, it makes sense - considering how those things will be used - to just create one of everything for each project
- the moment you create a new project, a bucket, a space for dashboards, somewhere to save time series etc. are all available
- unused S3 bucket / ... is virtually free
--
=> per project
---
layout: false

# Example Scopes

```
group:manage-members
```
<br>
???
- in the context of a group that bundles users, managing who is a member of the group is meaningful scope
--

```
project:bucket-write
project:clickhouse-read
...
```
<br>
???
- for a project, the different things created in all the components should be configurable separately => service provider might need to upload files but has no business modifying the ClickHouse tables for the project
--

```
project:admin
tenant:admin
group:admin
...
```
<br>
???
- includes everything, this is the scope that allows everything for the whole subtree
- (optional) granting `project:admin` on a tenant resource is possible and equivalent to granting it on all projects within, does not affect anything else like groups
--

```
project:view
tenant:view
group:view
...
```
<br>
???
- in some contexts even the list of project names / customers / user groups should not be visible to everyone, for those that are not granted `:view` scope it looks as if it doesn't exists
---
layout: true

# Keycloak ↔ Components

```{incgraph name='kccomp', height=200}
digraph {
  rankdir=LR
  k[label=Keycloak]
  t[label="Token Claims\nRun Ahead"] #1
  c[label="Superset\nGrafana\nPrometheus\nClickHouse"]
  k -> t -> c
}
```

---
`r kccomp(0)`
???
- we have the permissions in Keycloak and the Component needs to make an authorization decision
- how is authorization information communicated implementation-wise?
- ideally every software we want to use supports Keycloak Authorization Services on its own
- while some does, it's the exception rather than the rule
- one approach would be to add that generic support for everything => unfeasible to impossible
- alternative
  - talk to every component in a way it understands
  - use suitable existing mechanisms to communicate enough information to model the special cases we need
---
`r kccomp(1)`
???
- either include information in token
- or - while user is sent to Keycloak for authentication
  - pause token creation
  - talk to component over its API
  - ensure user exists and access is configured exactly like we need it
  - when token creation completes authorization information is already reconciled in component
  - after completion of authentication in component everything looks as expected
--

```json
{
  ...
  "https://aws.amazon.com/tags": {
    "principal_tags": {
      "bucket": ["tenant1.project1", "tenant1.project2"],
    }
  },
  ...
}
```
???
- example S3 (compatible, Ceph)
- this is what S3 (compatible) understands
- by using that mechanism we can integrate unmodified Ceph for buckets
---
layout: false

# Configuration
- necessary expressivity
- regular use simpler
???
- the underlying model needs to be expressive enough to cover the use cases that can reasonably arise
- for most situations the flexibility is not necessary
- => the user interface and offer simple ways to configure the most common cases, such as allowing full access to a project
---
class: inverse, center, middle

# Q(+A?)
